Longer term All-Cause and Cardiovascular Mortality with Intensive Blood Pressure Control

### 2021-12-31 doc/manuscript.Rmd
Initial commit

---
title: "Longer term All-Cause and Cardiovascular Mortality with Intensive Blood Pressure Control"
bibliography: refs.bib
csl: jacc.csl
always_allow_html: true
output: 
  officedown::rdocx_document:
    reference_docx: style_manuscript_times_new_roman.docx
---

@@

tar_load(names = c(recoders,
                   base_viz_inc,
                   base_viz_eff,
                   long_viz_forest,
                   base_ds, 
                   long_ds,
                   long_inf,
                   base_inf,
                   ndi_baseline))

@@

@@

@@

@@

```{r}
# in case you want to make something landscape,
# wrap it in this code
#<!---BLOCK_LANDSCAPE_START--->
#<!---BLOCK_LANDSCAPE_STOP--->
# remove the # signs - you will be placing this in the markdown text
```

\newpage
__Table 1__: All-cause mortality by treatment group and subgroup

tbl_data <- long_ds |>
  left_join(base_ds) |>
  mutate(
    .n = .("{n_acm} / {n_obs}"),
    .i = .("{incidence_est}\n({incidence_lwr}, {incidence_upr})")
  ) |>
  select(-starts_with("incidence_"),
         -starts_with("n_"),
         -starts_with("fup_")) |> 
  pivot_wider(names_from = c(treatment, trial_phase),
              values_from = c(.n, .i)) |> 
  left_join(tbl_inf) |> 
  mutate(
    variable = replace(variable, group == variable, NA_character_),
    variable = recode(variable, !!!recoders$variable),
    group = recode(group, !!!recoders$group)
  )

tbl_object <- tbl_data |> 
  select(variable, group, 
         ends_with('trial'), 
         ends_with('cohort')) |> 
  select(variable,
         group,
         # .n_Standard_trial,
         # .n_Intensive_trial,
         .i_Standard_trial,
         .i_Intensive_trial,
         .h_trial,
         # .n_Standard_cohort,
         # .n_Intensive_cohort,
         .i_Standard_cohort,
         .i_Intensive_cohort,
         .h_cohort) |> 
  as_grouped_data(groups = 'variable') |> 
  drop_na_rows() |> 
  as_flextable(hide_grouplabel = TRUE) |> 
  add_header_row(
    values = c(" ", 
               # "N Events / N Total", 
               "Incidence (95% CI)", 
               "Hazard Ratio\n(95% CI)",
               # "N Events / N Total", 
               "Incidence (95% CI)",
               "Hazard Ratio\n(95% CI)"),
    colwidths = c(1, 2, 1, 2, 1)
  ) |> 
  add_header_row(
    values = c(" ", 
               "Trial Follow-up Through Close-out Visits",
               "Post-trial Follow-up"),
    colwidths = c(1, 3, 3)
  ) |> 
  set_header_labels(
    group = " ",
    # .n_Standard_trial   = "Standard",
    # .n_Intensive_trial  = "Intensive",
    .i_Standard_trial   = "Standard",
    .i_Intensive_trial  = "Intensive",
    .h_trial            = "Hazard Ratio\n(95% CI)",
    # .n_Standard_cohort  = "Standard",
    # .n_Intensive_cohort = "Intensive",
    .i_Standard_cohort  = "Standard",
    .i_Intensive_cohort = "Intensive",
    .h_cohort           = "Hazard Ratio\n(95% CI)"
  ) |> 
  merge_v(part = 'header') |>
  theme_box() |> 
  padding(j = 1, i = ~ !is.na(group), padding.left = 15, part = 'body') |> 
  bg(i = ~ is.na(group), bg = 'grey80') |> 
  italic(i = ~ is.na(group), italic = TRUE) |> 
  width(width = 0.95) |> 
  width(j=1, width = 1.5) |> 
  align(align = 'center', part = 'all') |> 
  align(j = 1, align = 'left', part = 'all') |> 
  fontsize(size = 12, part = 'all') |> 
  font(fontname = 'Times New Roman', part = 'all')

@@

__Figure 1__: (A) Cumulative incidence of all-cause mortality by treatment group. (B) Time-dependent effect of being randomized to intensive treatment on all-cause mortality.

@@

\newpage
__Figure 2__: (A) Cumulative incidence of cardiovascular and non-cardiovascular mortality by treatment group. (B) Time-dependent effect of being randomized to intensive treatment on cardiovascular mortality.

@@

\newpage
__Figure 3__: Hazard ratio for all-cause mortality for participants randomized to intensive versus standard treatment, overall and in subgroups.

@@

long_viz_forest

@@

We have a total of `r nrow(base_viz_inc)` plots showing all-cause mortality and mortality split into cardiovascular and non-cardiovascular causes. They are automated enough to dynamically set text and axis values that will look reasonable. To show what I mean, I've added the figures for adults 75 years or older, who have a much higher incidence.

\newpage
__Figure S1__: Cumulative incidence of all-cause mortality among participants 75 years or older.

# base_viz_inc$acm[[3]]

\newpage
__Figure S2__: Cumulative incidence of cardiovascular and non-cardiovascular mortality among participants 75 years or older.

# base_viz_inc$cvd[[3]]

@@

>>>>

---
title: "Longer term All-Cause and Cardiovascular Mortality with Intensive Blood Pressure Control"
bibliography: refs.bib
csl: lancet.csl
always_allow_html: true
output: 
  officedown::rdocx_document:
    reference_docx: style_manuscript_times_new_roman.docx
---

@@

tar_load(names = c(recoders,
                   base_viz_inc,
                   base_viz_eff,
                   long_viz_mort,
                   long_viz_cvd,
                   base_ds, 
                   long_ds,
                   long_inf,
                   base_inf,
                   ndi_baseline))

@@

<!-- END OF ABSTRACT -->

@@

___Statistical Analysis___: The effect of randomization to intensive treatment was estimated as a function of time using two approaches. The first approach split the follow-up of study participants into non-overlapping trial and cohort phases, then estimated the effect of randomization to intensive treatment in each phase [@zhang_time_varying_2018]. The second approach estimated the cumulative time-dependent effect of randomization to intensive treatment [@martinussen2007dynamic]. While the phase-specific estimate of the treatment effect treats the trial and cohort phases independently, the cumulative estimate enforces continuity of the treatment effect through time so that the effect of treatment during the trial phase is not 'forgotten' during the cohort phase. All analyses accounted for correlation within study sites [@lin_robust_1989], and analyses of cardiovascular mortality accounted for the competing risk of non-cardiovascular mortality [@scheike_flexible_2008].

@@

In both treatment groups, median follow-up time was `r ap_b_o$fup_est` years. A total of `r ap_b_i$n_acm` and `r ap_b_s$n_acm` all-cause mortality events occurred among participants randomized to intensive and standard treatment, respectively (__Table 1__). The hazard ratio (HR) for all-cause mortality among participants randomized to intensive versus standard treatment was `r with(ap_l_o_t, .("{mort_est} (95% confidence interval [CI] {mort_lwr}, {mort_upr})"))` during the trial phase and `r with(ap_l_o_c, .("{mort_est} (95% CI {mort_lwr}, {mort_upr})"))` during the cohort phase. The cumulative time-varying effect of intensive versus standard treatment indicated lower risk for all-cause mortality during the trial phase and was attenuated during the cohort phase (__Figure 1__).

@@

@@

__Table 1__: All-cause mortality by treatment group and subgroup

tbl_data <- long_ds |>
  mutate(
    across(starts_with("n_"), as.integer),
    .n = .("{n_acm} / {n_obs}"),
    .i = .("{inc_mort_est}\n({inc_mort_lwr}, {inc_mort_upr})")
  ) |>
  select(-starts_with("inc_"),
         -starts_with("n_")) |> 
  pivot_wider(names_from = c(treatment, trial_phase),
              values_from = c(.n, .i)) |> 
  left_join(tbl_inf) |> 
  mutate(
    variable = replace(variable, group == variable, NA_character_),
    variable = recode(variable, !!!recoders$variable),
    group = recode(group, !!!recoders$group)
  )

tbl_object <- tbl_data |> 
  select(variable,
         group,
         .n_Standard_trial,
         .n_Intensive_trial,
         .i_Standard_trial,
         .i_Intensive_trial,
         .h_trial,
         .n_Standard_cohort,
         .n_Intensive_cohort,
         .i_Standard_cohort,
         .i_Intensive_cohort,
         .h_cohort) |> 
  as_grouped_data(groups = 'variable') |> 
  drop_na_rows() |> 
  as_flextable(hide_grouplabel = TRUE) |> 
  add_header_row(
    values = c(" ", 
               "N Events / N Total", 
               "Incidence\n(95% CI)", 
               "Hazard Ratio\n(95% CI)",
               "N Events / N Total", 
               "Incidence\n(95% CI)",
               "Hazard Ratio\n(95% CI)"),
    colwidths = c(1, 2, 2, 1, 2, 2, 1)
  ) |> 
  add_header_row(
    values = c(" ", 
               "Trial Follow-up Through Close-out Visits",
               "Post-trial Follow-up"),
    colwidths = c(1, 5, 5)
  ) |> 
  set_header_labels(
    group = " ",
    .n_Standard_trial   = "Standard",
    .n_Intensive_trial  = "Intensive",
    .i_Standard_trial   = "Standard",
    .i_Intensive_trial  = "Intensive",
    .h_trial            = "Hazard Ratio\n(95% CI)",
    .n_Standard_cohort  = "Standard",
    .n_Intensive_cohort = "Intensive",
    .i_Standard_cohort  = "Standard",
    .i_Intensive_cohort = "Intensive",
    .h_cohort           = "Hazard Ratio\n(95% CI)"
  ) |> 
  merge_v(part = 'header') |>
  theme_box() |> 
  padding(j = 1, i = ~ !is.na(group), padding.left = 15, part = 'body') |> 
  bg(i = ~ is.na(group), bg = 'grey80') |> 
  italic(i = ~ is.na(group), italic = TRUE) |> 
  width(width = 0.95) |> 
  width(j=1, width = 1.5) |> 
  align(align = 'center', part = 'all') |> 
  align(j = 1, align = 'left', part = 'all') |> 
  fontsize(size = 12, part = 'all') |> 
  font(fontname = 'Times New Roman', part = 'all')

\newpage
__Table 2__: Cardiovascular and non-cardiovascular mortality by treatment group and subgroup

tbl_inf <- long_inf |> 
  mutate(cvd = .("{cvd_est}\n({cvd_lwr}, {cvd_upr})"),
         non = .("{other_est}\n({other_lwr}, {other_upr})")) |> 
  select(variable,
         group, 
         trial_phase,
         cvd,
         non) |> 
  pivot_wider(names_from = trial_phase,
              values_from = c(cvd, non),
              names_glue = ".h_{.value}_{trial_phase}")

tbl_data <- long_ds |>
  mutate(
    across(starts_with("n_"), as.integer),
    .n = .("{n_cvd} / {n_other} / {n_obs}")
  ) |>
  select(-starts_with("inc_"),
         -starts_with("n_")) |> 
  pivot_wider(names_from = c(treatment, trial_phase),
              values_from = c(.n),
              names_prefix = '.n_') |> 
  left_join(tbl_inf) |> 
  mutate(
    variable = replace(variable, group == variable, NA_character_),
    variable = recode(variable, !!!recoders$variable),
    group = recode(group, !!!recoders$group)
  )

tbl_object <- tbl_data |> 
  select(variable,
         group,
         .n_Standard_trial,
         .n_Intensive_trial,
         .h_cvd_trial,
         .h_non_trial,
         .n_Standard_cohort,
         .n_Intensive_cohort,
         .h_cvd_cohort,
         .h_non_cohort) |> 
  as_grouped_data(groups = 'variable') |> 
  drop_na_rows() |> 
  as_flextable(hide_grouplabel = TRUE) |> 
  add_header_row(
    values = c(" ", 
               "N\nCVD / non-CVD / Total", 
               "Hazard Ratio\n(95% CI)",
               "N\nCVD / non-CVD / Total", 
               "Hazard Ratio\n(95% CI)"),
    colwidths = c(1, 2, 2, 2, 2)
  ) |> 
  add_header_row(
    values = c(" ", 
               "Trial Follow-up Through Close-out Visits",
               "Post-trial Follow-up"),
    colwidths = c(1, 4, 4)
  ) |> 
  set_header_labels(
    group = " ",
    .n_Standard_trial= "Standard",
    .n_Intensive_trial= "Intensive",
    .h_cvd_trial = "CVD",
    .h_non_trial = "Non-CVD",
    .n_Standard_cohort = "Standard",
    .n_Intensive_cohort = "Intensive",
    .h_cvd_cohort = "CVD",
    .h_non_cohort = "Non-CVD"
  ) |> 
  merge_v(part = 'header') |>
  theme_box() |> 
  padding(j = 1, i = ~ !is.na(group), padding.left = 15, part = 'body') |> 
  bg(i = ~ is.na(group), bg = 'grey80') |> 
  italic(i = ~ is.na(group), italic = TRUE) |> 
  width(width = 0.95) |> 
  width(j=1, width = 1.5) |> 
  align(align = 'center', part = 'all') |> 
  align(j = 1, align = 'left', part = 'all') |> 
  fontsize(size = 12, part = 'all') |> 
  font(fontname = 'Times New Roman', part = 'all')

\newpage
__Figure 1__: (A) Cumulative incidence of all-cause mortality by treatment group. (B) Time-dependent effect of randomization to intensive treatment for all-cause mortality.

@@

\newpage
__Figure 2__: (A) Cumulative incidence of cardiovascular and non-cardiovascular mortality by treatment group. (B) Time-dependent effect of randomization to intensive treatment for cardiovascular mortality.

@@

\newpage
__Figure 3__: Mean systolic blood pressure over time by treatment group. Shaded areas indicate a 95% confidence interval for the mean.

@@

knitr::include_graphics(
  path = 'Z:/bjaeger/sprint-mortality-ndi/fig/SPRINT_EHR_SBP_LongTerm_Comparison_122721.png', 
  dpi = 600
)

\newpage
__Figure 4__: All-cause mortality hazard ratio for participants randomized to intensive versus standard treatment.

long_viz_mort

\newpage
__Figure 5__: Cardiovascular mortality hazard ratio for participants randomized to intensive versus standard treatment.

long_viz_cvd

@@

To be decided.

@@

# REFERENCES
